---
title: "Models"
author: "Melissa Ban"
---
```{r}
#| echo: false
#| message: false
library(tidyverse)
library(usethis)
library(primer.data)
library(rstanarm)
```

## Hypothesis 

Null hypothesis: there is no correlation between patients’ initial treatment ages and the percent improvement in SE (%).

Alternative hypothesis: there is a correlation between patients’ initial treatment ages and the percent improvement in SE (%).

## Calculations


Initial SE (D): $$SE_i = SPH_i + 1/2CYL_i$$


## Processed data table

```{r}
#| echo: false
#| message: false
raw_data <- readxl::read_excel("Raw data.xlsx") |>
  mutate(gender = c("F", "F", "M", "M", "F", "F", 
                    "F", "F", "F", "F", "F", "F", 
                    "M", "M", "M", "M", "M", "F", 
                    "F", "F", "M", "M", "F", "M", 
                    "M", "M", "M", "M", "F", "F", 
                    "F", "M", "F", "M")) 
  
x <- raw_data |>
    mutate(SE_i = SPH_i + 1/2 *CYL_i,
            SE_f = SPH_f + 1/2 *CYL_f,
            SE_d = SE_f - SE_i,
            SE_improve = abs(SE_d/SE_i)) |>
  summarise(SE_avg_imp = abs(mean(SE_improve)), .by = Age)
  
x
```

## Correlation Model

```{r}
#| echo: false
#| message: false

raw_data <- readxl::read_excel("Raw data.xlsx") |>
  mutate(gender = c("F", "F", "M", "M", "F", "F", 
                    "F", "F", "F", "F", "F", "F", 
                    "M", "M", "M", "M", "M", "F", 
                    "F", "F", "M", "M", "F", "M", 
                    "M", "M", "M", "M", "F", "F", 
                    "F", "M", "F", "M")) 
  
x <- raw_data |>
    mutate(SE_i = SPH_i + 1/2 *CYL_i,
            SE_f = SPH_f + 1/2 *CYL_f,
            SE_d = SE_f - SE_i,
            SE_improve = abs(SE_d/SE_i)) 

model <- lm(x$SE_improve~x$Age)

summary(model)

fitstan <- stan_glm(x, formula = SE_improve ~ gender, family = gaussian, refresh = 0, iter = 8000, seed = 100)


newobs = tibble(gender = c("F", "M"))

pp <- posterior_epred(fitstan, newdata = newobs) |>
  as.tibble() |>
  pivot_longer(`1`:`2`,
               names_to = "gender",
               values_to = "values") |>
  mutate(gender = if_else(gender == "1", "Female", "Male"))

pp |>
  ggplot(aes(x = values, fill = gender)) + 
  geom_histogram(aes(y = after_stat(count/sum(count))),
                 position = "identity", alpha = 0.6, bins = 100) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 0.1))

plot(x$Age, x$"SE_improve")

```

Model: 
$$y_i = \alpha*Age + \beta*Gender$$
## Pearson

When considering the relationship between the two factors, Pearson’s correlation coefficient and Spearman's ranking order correlation coefficient are the most common tools. Pearson’s test is best used when there are no outliers in the data (Kent State University, 2021), whereas Spearman’s test is most accurately used when the data has significant outliers. According to Table 5, there are no outliers in this set of data. Therefore, Pearson’s test is chosen.

```{r}
#| echo: false
#| message: false

raw_data <- readxl::read_excel("/Users/melissashedgehog/Desktop/Projects/OrthoK Lenses Effect/Raw data.xlsx")

x <- raw_data |>
    mutate(SE_i = SPH_i + 1/2 *CYL_i,
            SE_f = SPH_f + 1/2 *CYL_f,
            SE_d = SE_f - SE_i,
            SE_improve = (SE_d/SE_i))

model <- lm(x$SE_d~x$Age)

summary(model)

ggplot(x, aes(x = Age, y = SE_d)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(title = "Relationship between patients’ initial treatment ages and \n the average percent improvement of SE (%).",
       subtitle = "For age groups of 8, 9, 10, 11, 12, 13, and 14",
       x = "Patients' Initial Treatment Ages",
       y = "Average Percent Improvement of Spherical Equivalence (%)") +
  theme_minimal()
```

## Pearson

When considering the relationship between the two factors, Pearson’s correlation coefficient and Spearman's ranking order correlation coefficient are the most common tools. Pearson’s test is best used when there are no outliers in the data (Kent State University, 2021), whereas Spearman’s test is most accurately used when the data has significant outliers. According to Table 5, there are no outliers in this set of data. Therefore, Pearson’s test is chosen.

```{r}
#| echo: false
#| message: false

correlation <- cor.test(x$Age, x$`SE_improve`, method = "pearson")

cor.test(x$Age, x$SE_d)
```

If the coefficient > 0, then there is a positive correlation. Which is, by increasing one value, the other value also increases. If the coefficient < 0, then there is a negative correlation, meaning by increasing one value, the additional value decreases. 

Here, the r-value is 0.4902269, which is greater than 0. Since this value is between 0 and 0.5, it is considered a weak positive correlation. 

Although it is suggested that there is a weak positive correlation between two variables, we still nedd to examine its statistical significance, that is, whether the relationship with this sample exists in the population (University of Connecticut, 2009). A significance level (α) is chosen to be 0.05, meaning at least 95% times the relationship exhibited in this sample also exists in the populations (University of Connecticut, 2009). Since the p-value is 0.003249, which is smaller than 0.05, the relationship is statistically significant. Therefore, the alternative hypothesis is accepted.
